---
title: "Homework 2"
author: "Jakob Serlier"
date: "Assigned: 29 Oct 2019"
output: 
  html_document:
    theme: paper
    highlight: tango
---

## This homework is due by **18:00 on Wednesday, 6 November**.  To complete this assignment, follow these steps:

1. Download the `homework2.Rmd` file from Canvas.
2. Open `homework2.Rmd` in RStudio.
3. Replace the "Your Name Here" text in the `author:` field with your own name.
4. Supply your solutions to the homework by editing `homework1.Rmd`.
5. When you have completed the homework and have **checked** that your code both runs in the Console and knits correctly when you click `Knit HTML`, rename the R Markdown file to `homework1_YourNameHere.Rmd`, and submit on Canvas.  (YourNameHere should be changed to your own name.)

## Homework tips:

1. Instead of sending code line-by-line with `<ctrl-enter>`, you can send entire code chunks, and even run all of the code chunks in your .Rmd file. Look under the <Chunks> menu of the Source panel.
2. Run your code in the Console and Knit HTML frequently to check for errors.
3. You may find it easier to solve a problem by interacting only with the Console at first. 

## Homework 1 outline

## Problem 1: Simple Boolean operations

> Tip: To check that your answer makes sense, you may want to try it out in the console with various choices of values for the variable `x`.   

## (a) Checking equality.

Given a variable `x`, write a Boolean expression that evaluates to `TRUE` if the variable `x` is equal to `94842` (the numeric value).  

```{r question 1}
# Insert your Boolean expression here
```

## (b) Checking inequality.

Given a variable `x`, write a Boolean expression that evaluates to `TRUE` if the variable `x` is *not* `NA` (i.e., is not missing).

```{r}
# Insert your Boolean expression here
```

## (c) Checking if a number is in a given range. 

Given a (possibly negative) number `x`, write a Boolean expression that returns `TRUE` if and only if `x` is smaller than `-12` or bigger than `29`.  

```{r}
# Insert your Boolean expression here
```

## Problem 2: Vector Boolean operations

## (a) R has two kinds of Boolean operators implemented, single (`&`, `|`) and double (`&&`, `||`).  

One of these operators takes advantage of something called *lazy evaluation* while the other does not.  They also don't behave the same way when applied to *vectors*.  

Read the help file (`help("||")`) and construct some examples to help figure out how the two behave.  

To help you get started, try out the following two examples in your console:

```{r eval = FALSE}
# Example:  The variable y.prob2a is not defined.  
# (Do not define it!)
# What happens when you run this code?
x.prob2a <- 5
(x.prob2a < 10) | (y.prob2a > 2)
(x.prob2a < 10) || (y.prob2a > 2)
```

```{r}
# Define vectors
x.prob2a.vec <- c(TRUE, FALSE, FALSE)
y.prob2a.vec <- c(TRUE, TRUE, FALSE)

# Apply various Boolean operations to see what happens
x.prob2a.vec & y.prob2a.vec
x.prob2a.vec && y.prob2a.vec
x.prob2a.vec | y.prob2a.vec
x.prob2a.vec || y.prob2a.vec
```

Can you explain what's happening?  Write up a brief explanation below.

<font color="#336600">

**Replace this text with your explanation.**

</font>

## (b) Using `all()`

Two people were asked to give their preferences between two options: [Facebook, Twitter], [Firefox, Chrome], [Mac, PC], [Summer, Winter].  Their results are given below.

```{r}
alice.prefs <- c("Twitter", "Chrome", "Mac", "Summer")
bob.prefs <- c("Facebook", "Chrome", "PC", "Summer")
```

Use the `all()` function to determine if the two people have identical preferences. (Your code should ouput a single Boolean value, either `TRUE` or `FALSE`)

```{r}
# Edit me
```

## (c) Using `any()` 

Use the `any()` function to determine if the two people have any preferences in common.  (Your code should output a single Boolean value, either `TRUE` or `FALSE`)
```{r}
# Edit me
```

## (d) Missing values.

Let `age` be the vector defined below.

```{r}
age <- c(18, NA, 25, 71, NA, 45, NA, NA, 18)
```

Write a Boolean expression that checks whether each entry of `age` is missing (recall missing values are denoted by `NA`).  Your expression should return a Boolean vector having the same length as `age`.

```{r}
# Edit me
```

## Problem 3: Referencing vector elements

## (a)  `which()` practice

Write code that returns the indexes of `age` that are missing.

```{r}
# Edit me
```

## (b) Getting non-missing values

Write code that uses negative indexes and your solution from (a) to return only the values of `age` that are *not* missing. (i.e., your code should result in a vector with elements: 18, 25, 71, 45, 18) 


```{r}
# Edit me
```

## (c)  A more direct way of getting non-missing values

Using the negation operator `!` and the `is.na()` function, write an expression that returns only the values of `age` that are *not* missing.

```{r}
# Edit me
```

## (d) More `which()` practice

For the next three problems we will go back to the `cars` data set.

```{r}
speed <- cars$speed
dist <- cars$dist
```

Write code to figure out which cars had a stopping distance of 15 feet or more.

```{r}
# Edit me
```

## (e) `which.min`, `which.max` practice

Use the `which.min()` function to figure out which car had the *shortest* stopping distance. (Your code should return the car's index.)

```{r}
# Edit me
```

## (f) More practice 

Use the `which.max()` function to figure out the *speed* of the car that had the *longest* stopping distance.  (Your code should return the car's speed.)

```{r}
# Edit me
```

## Problem 4: Data frame basics

## (a) Importing data.

In Lecture 2 we saw how to use the `read.table()` function to import the survey data.  Now we'll use a different function.  Use the `read.csv()` function to import the survey data into a variable called `survey`.

```{r}
# Edit me
```

**Tip**: The data file is available on Canvas.  You may download the file to a directory on your machine then import it.

## (b) `$` notation

Use the `$` operator to select the TVhours column from the `survey` data

```{r}
# Edit me
```

## (c) [,] notation

Repeat part (b) using `[,]` notation.  i.e., Use `[,]` notation to select the TVhours column from the `survey` data by name (i.e., obtain this column by using the name "TVhours" instead of using the column number)

```{r}
# Edit me
```

## (d) [[]] notation

Repeat part (c) with [[]] notation.

```{r}
# Edit me
```

## (e) [] notation

Repeat part (d), but this time using single blackets (`[ ]`) notation.  

(Observe that this returns a new single-column *data frame*, not just a vector.) 

```{r}
# Edit me
```

## (f) `subset()` practice

Use the `subset()` function to select all the survey data on Program and OperatingSystem for respondents whose Rexperience is "Never used" *or* who watched 5 or more hours of TV last week.

```{r}
# Edit me
```

## Problem 5: Data summaries and inline code practice.

## (a) Bar graph

Create a bar graph of respondents' Rexperience.

```{r, fig.align='center', fig.width=7, fig.height=4}
# Edit me
```

## (b) Inline code practice

Replace all occurrences of ???? in the paragraph below with an inline code chunk supplying the appropriate information.

> Of the ???? survey respondents, ???? were NOT from the MISM program.  We found that ????% of the all students in the class use the Mac OS X operating system.  ????% of of MISM students report having Basic competence in R.